---
title: "Assignment 3"
author: "Taelor Malcolm"
date: "9/20/2020"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
library(maps)
library(viridis)
```
# Loading and Transforming Data 

```{r load data}
ncounties <- st_read("https://opendata.arcgis.com/datasets/53ca7db14b8f4a9193c1883247886459_67.kml", 
                  quiet = TRUE) 

greenspace <- st_read("https://opendata.arcgis.com/datasets/090b68b1e926485fbb1840b5c2e710ac_94.kml", 
                  quiet = TRUE)

gardens <- st_read("https://opendata.arcgis.com/datasets/20847c0908834a6bb1e3474275985640_123.kml", 
                  quiet = TRUE)

hospitals <- st_read("https://opendata.arcgis.com/datasets/e63a182a16a24b0680f940ed6e7b80f4_10.kml", 
                  quiet = TRUE)

```

### Point Data 

A.Medical facilities  
B.Gardens, farms, and orchards 

### Polygon Data

C. Counties in Atlanta  
D. Greenspace

```{r}
GA_state_plane <- "+proj=tmerc +lat_0=30 +lon_0=-82.16666666666667 +k=0.9999 +x_0=200000 +y_0=0 +ellps=GRS80 +units=m +no_defs "

#I chose the Georgia StatePlane East projection for this assignment. 

ncounties <- ncounties %>%
  st_transform(GA_state_plane)

greenspace <- greenspace %>%
  st_transform(GA_state_plane)

gardens <- gardens %>%
  st_transform(GA_state_plane)

hospitals <- hospitals %>%
  st_transform(GA_state_plane)

```

# Points within a polygon
## Figure 1
The number and proportion of garden points within 700 meters of a hospital.
```{r}
garden_buffer <- st_buffer(gardens, dist = 700) %>%
  st_union()

ggplot(garden_buffer) +
  geom_sf() +
  theme_map()
```
```{r}
hospitals_garden <- hospitals[garden_buffer,]
  
ggplot(garden_buffer) +
  geom_sf() +
  geom_sf(data = hospitals_garden, 
          color = "forestgreen", 
          size = 1.5) +
  theme_map()
```
```{r}
hospitals <- hospitals %>%
  st_join(hospitals_garden) %>%
  mutate(by_garden = !is.na(Name.y))


n_gardens_hospitals <- sum(hospitals$by_garden)

n_gardens_hospitals
```
There are 12 hospitals within 700 meters of a garden. 
```{r}
n_hospitals <- length(hospitals$by_garden)

pct_gardens_hospitals <- n_gardens_hospitals / n_hospitals

pct_gardens_hospitals
```
About 27% of all hospitals in Atlanta are within 700 meters of a garden.

# Average Distance 
## Figure 2
This is the average distance between gardens and medical facilities. 
```{r}
gardens <- gardens %>%
  mutate(hospital_dist = st_nn(gardens, hospitals, 
                           returnDist = TRUE)$dist) %>%
  mutate(hospital_dist = as.numeric(hospital_dist))


```

```{r}
avg_hospital_dist <- mean(gardens$hospital_dist)

avg_hospital_dist
```
```{r}
right_side <- st_bbox(gardens)$xmax
left_side  <- st_bbox(gardens)$xmin
top_side <- st_bbox(gardens)$ymax
bottom_side <- st_bbox(gardens)$ymin


ggplot(greenspace) +
  geom_sf(fill = "olivedrab", color = NA) +
  geom_sf(data = gardens, size = 0.1,
          aes(color = hospital_dist)) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_color_viridis_c(name = 
                          "Atlanta gardens \nby distance to a hospital") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = right_side - 50000, 
           y = bottom_side + 9000, 
           label = paste("On average, a garden in Atlanta \nis ", 
                         prettyNum(avg_hospital_dist, digits = 0),
                         " meters from a hospital.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "gray95"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

# Number of Points
## Figure 3
The number and proportion of county polygons containing hospitals.

```{r}
ncounties <- ncounties %>%
  mutate(num_hospitals = lengths(st_covers(ncounties, hospitals)))

ggplot(ncounties) +
  geom_sf(color = NA, 
          aes(fill = num_hospitals)) +
  scale_fill_viridis_c(name = "Atlanta counties\nby number of hospitals",
                       breaks = breaks <- seq(0, 20, by = 5),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "hospitals")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray")) 
```
 
## Figure 4
The number and proportion of county polygons containing garden points.
```{r}
ncounties <- ncounties %>%
  mutate(num_gardens = lengths(st_covers(ncounties, gardens)))

ggplot(ncounties) +
  geom_sf(color = NA, 
          aes(fill = num_gardens)) +
  scale_fill_viridis_c(name = "Atlanta counties\nby number of gardens",
                       breaks = breaks <- seq(0, 250, by = 50),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "gardens"), 
                       option = "plasma") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray")) 
```

# Overlapping Polygons
## Figure 5
Overlapping county and greenspace polygons. 
```{r}
ncounties <- ncounties %>%
  mutate(num_greenspace = lengths(st_overlaps(ncounties, greenspace))) %>%
  mutate(has_greenspace = num_greenspace > 0)
```

```{r}
n_greenspace_ncounties <- sum(ncounties$has_greenspace)

n_greenspace_ncounties
```

```{r}
left_side  <- st_bbox(greenspace)$xmin
top_side <- st_bbox(greenspace)$ymax

ggplot(greenspace) +
  geom_sf(fill = "olivedrab", color = NA) +
  geom_sf(data = ncounties,
          aes(fill = has_greenspace)) +
  scale_fill_manual(values = c("ivory", "olivedrab"),
          name = "Counties in Atlanta\nby presence of greenspace", 
          labels = c("County without\noverlappingg greenspace",
                     "County with \noverlapping greenspace")) +
  annotation_scale(location = "br") +
  annotate(geom = "text", x = left_side, 
           y = top_side -1000, 
           label = paste(n_greenspace_ncounties ,
                         "of Atlanta's", 
                         length(ncounties$Name),
                         "counties \noverlap with", 
                         "greenspace."),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "gray90"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "lightgray"))
```

# Density
## Figure 6
The average density of garden points in each county polygon
```{r}
ncounties <- ncounties %>%
  mutate(area = set_units(st_area(ncounties), km^2)) %>%
  mutate(gardens_dens = as.numeric(num_gardens / area))

ggplot(ncounties) +
  geom_sf(color = NA, 
          aes(fill = gardens_dens)) +
    scale_fill_viridis_c(name = 
                           "Counties in Atlanta\nby garden density",
                       breaks = breaks <- seq(0, 1, by = .05),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "gardens per square km"), 
                       option = "inferno") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_minimal()) +
  theme_map() +
  theme(legend.position = "right")

```



